\vspace{-1cm} Model simplicity breeds contempt: using simple models to answer basic questions on species' distributions

### 2021-02-12 manuscript/manuscript.tex
Trying to do anything to get this to converge

One of the central goals of ecology is to understand the ways species are distributed across space and time (ref). As a result, ecologists have developed multiple distribution models over the last two decades, to try to untangle the factors that play a role in defining such distributions (ref). These models estimate species' realized niches using several covariates, including environmental variables (ref), species ecological traits' (ref) and phylogenetic relations (ref). For example, ... More recently, some of the focus have shifted towards approaches that estimate and account for biotic factors, such as competitive or facilitative relationships between species. The idea is that by understanding how such biotic and abiotic factors shape species' distributions, we can gain a mechanistic understanding on how ecological communities are established and change over time. However, while some of this approaches can increase the predictive performance of distribution models (ref), the nature of some of the estimates have been theoretically questioned (ref). This was parsimoniously illustrated by \citealt{bla bla}, who ...

>>>>

One of the central goals of ecology is to understand the ways species are distributed across space and time (ref). As a result, ecologists have developed multiple distribution models over the last two decades, to try to untangle the factors that play a role in defining such distributions (ref). These models estimate species' realized niches using several covariates, including environmental variables (ref), species ecological traits' (ref) and phylogenetic relations (ref). And more recently, some of the focus have shifted towards approaches that estimate and account for biotic factors, such as competitive or facilitative relationships between species (ref). The idea is that by understanding how such biotic and abiotic factors shape species' distributions, we can gain a mechanistic understanding on how ecological communities are established and change over time. However, while some of this approaches can increase the predictive performance of distribution models (ref), the nature of some of the estimates have been theoretically questioned (ref). This was recently best parsimoniously illustrated by \citealt{bla bla}, who ...

While the value of gaining a mechanistic understanding of species' distributions is unquestioned (ref), a lot can be learned from shedding light on basic properties of species' realized niches. Indeed, there is no general agreement the shape of species distributions. While many ecological textbooks (Begon et al., 1990, Giller, 1984, Krebs, 1994) assume this to be unimodal and symmetric, some have warned that empirical distributions can take many different forms \citep{austin2002spatial}.

There is not an easy way to untangle the true shape of species' distributions, as this shape is likely to showcase idiosyncrasies at the species level and across systems. The aim of this work, it is not to answer these questions nor to provide a general approach that accommodates such idiosyncrasies. Instead, we want to use a model that is solely constrained by the empirical information that we truly have regarding a particular system, relaxing as much as possible the structural constrains of the statistical framework. Then, we want to use this model to answer basic aspects regarding the way systems of many species are distributed along an environmental gradient.
